---
title: "Harry Potter and regular expressions"
author: 'STOR 390'
output: html_document
---

Assignment 2 is due 2/28/17. You can find the raw .Rmd file at: [https://raw.githubusercontent.com/idc9/stor390/master/assignments/harry_potter/harry_potter.Rmd](https://raw.githubusercontent.com/idc9/stor390/master/assignments/harry_potter/harry_potter.Rmd).

The text of all 7 Harry Potter books is available online: [http://www.readfreeonline.net/Author/J._K._Rowling/Index.html](http://www.readfreeonline.net/Author/J._K._Rowling/Index.html) (very possible the website is now defunct). In this assignment you will use dplyr, ggplot and regular expressions to do an exploratory analysis of Harry Potter and the Philosopher's Stone.

Here are a couple examples of similar text analysis projects (that you will be able to do in a couple weeks!)

- [The Life-Changing Magic of Tidying Text](http://juliasilge.com/blog/Life-Changing-Magic/) by Julia Silge (yes [janeaustenr](https://github.com/juliasilge/janeaustenr) is an entire R package devoted to Jane Austen)

- [Harry Potter agression](https://github.com/andrewheiss/Harry-Potter-aggression) by Andrew Heiss


# Question 0
Set `eval=FALSE` for the chunk above and `eval=TRUE`for the chunk below and all test chunks. The text file comes with the Sakai announcement.
```{r, eval=T, message=F, warning=F}
# set up
library(tidyverse)
library(stringr)
text <- read_file('philosophers_stone.txt')
```

# Question 1
How many words are in the book? 

```{r}
text %>%
  str_count("\\w+") %>%
  sum()
```


# Question 2 

How many times are each of the following characters mentioned? Display the answer using an appropriate visualization. 

- Harry, Hermione, Ron, Neville, Dumbledore, Draco, Snape, Hagrid, McGonagall


*Hint*: the `map` function might be helpful (but not necessary).


```{r}
people <- c('Harry', 'Hermione', 'Ron', 'Neville', 'Dumbledore', 'Draco', 'Snape', 'Hagrid', 'McGonagall')
characters <- tibble(name = factor(people), count = str_count(text, people))
characters
ggplot(characters, aes(x = name, y = count, fill = name)) +
  geom_bar(stat = 'identity') +
  theme(axis.ticks.x=element_blank(),
        axis.text.x=element_blank())


```






# Question 3
Break the text into paragraphs; create a vector called `paragraphs` where each entry is a paragraph in the book.


```{r}
# assume paragraphs end with \\\r\\\n
paragraphs <- text %>%
  str_c() %>%
  str_split('\\r\\n') %>%
  unlist()
```


# Question 4
Write a function that can break the text up into paragraphs, sentences, or words. This is a preview of [what you'll be doing](http://tidytextmining.com/tidytext.html#the-unnest_tokens-function) in a couple weeks.

This function does not need to be perfect. For sentences, give one example where the function you wrote fails.

*Hint*: the function should probably have a if statement
```{r}
unnest_tokens <- function(text, token='words'){
  # splits a string into tokens
  # input
  # text is a string
  # token can be one of: words, paragraphs, sentences
  # output: a character vector
  delimiter <- ''
  if (token == 'words')
    delimiter <- '[^a-zA-Z\']+'
  else if (token == 'paragraphs')
    delimiter <- '\\r\\n'
  else if (token == 'sentences')
    delimiter <- '(?<=[.?!])[\\s\\r\\n]+'
  else return(NULL)
  text %>%
    str_split(delimiter) %>%
    unlist()
}
```


```{r test4, eval=F}
# TODO add more
# Test code for the grader -- you don't have to modify these
sum(paragraphs == unnest_tokens(text, 'paragraphs'))
```


# Question 5

Put the data into tidy format with one row per paragraph. 

- first remove all paragraphs length 0
- create a tibble called `paragraph_df` with one column `text` with the text of each paragraph (*hint*: you might need to use `as.character(paragraphs)`)
- add a new column `index` that gives the index of each paragraph
- **wighout** using dplyr add a column called `Harry` that counts the number of times Harry is referenced in each paragraph


```{r}
paragraph_df <- tibble(text = as.character(paragraphs), 
                       index = 1:length(paragraphs),
                       Harry = str_count(paragraphs, 'Harry'))
```

*Hint*: you can use question 2 to check your answer

# Question 6
Write a function called `reference_counter` that generalizes question 5 for any tidy text data frame and any list of words. 

*Hint*: do this **without** dplyr

```{r}

reference_counter <- function(text_df, word_list){
  
  # inputs
  # text_df is a tibble with a column called text
  # word_list is a vector of strings
  # for each word in word_list add a column to text_df counting
  # the number of times that word appears in each row of text df
  # does not modify the original text_df
  # do this WITHOUT using dplyr
  res <- text_df
  for (word in word_list)
    res[[word]] <- str_count(res$text, word)
  return(res)
}
```


```{r test6, eval=F}
# test code for grader
test_words <- c('Harry', 'Hagrid', 'wand')
test_df <- reference_counter(paragraph_df, test_words)
test_df %>% select(Harry, Hagrid, wand) %>% summarise_all(sum)
```



# Question 7 

Using the `reference_counter` function update `paragraph_df` to include columns counting the number of references to each characters from Q2 in each paragraph


```{r}
paragraph_df <- reference_counter(paragraph_df, people)
```



```{r test7, eval=F}
# test code for grader
paragraph_df[,people] %>% summarise_all(sum)
paragraph_df
```


# Question 8

Make a new data frame called `person_refs`  with three columns: person, num_refs, index. num_refs is the number of references each person gets in paragraph and index is the index of the paragraph. Limit this data frame to the following 5 characters: Harry, Hermione, Ron, Draco, Neville. 

*Hint*: use `gather`.

```{r}
person_refs <- paragraph_df %>%
  select(index, Harry, Hermione, Ron, Draco, Neville) %>%
  gather(Harry, Hermione, Ron, Draco, Neville,
         key = 'person', value = 'num_refs')
```


Make a bar plot showing the number of paragraphs that references each of the 5 characters

```{r}
ggplot(person_refs, aes(x = person, y = as.numeric(num_refs > 0))) +
  geom_bar(stat = 'identity')
```


Now we want to examine how characters evolve over "time." Plot the number of references vs. the paragraph index.

```{r}
ggplot(person_refs, aes(x = index, y = cumsum(num_refs), color = person)) + 
  geom_line()
```

In this question we are using paragraphs for "time windows." What are other "time windows" we could have used? What are some trade offs for these different choices.

We could have used lines, words or sentences. Lines or words would be more consistent as a method of "timekeeping" since the length is more consistent, but being smaller they would leave larger gaps in the data.

# Question 9
How often are Harry and Herminone referenced together? Plot the number of references per paragraph for Harry vs. Herminone.

- one plot using `geom_point`
- one plot using `geom_jitter` (use the width/height arguments of jitter to make the jitter plot look better)


```{r}
tmp <- paragraph_df %>%
  filter(Harry > 0 & Hermione > 0)
length(tmp$index)

ggplot(paragraph_df, aes(x = Harry, y = Hermione)) + 
  geom_point()
```


```{r}
ggplot(paragraph_df, aes(x = Harry, y = Hermione)) + 
  geom_jitter(width = .3, height = .3)
```


Why is the jitter plot better than a simple point plot?

When both characters have the same number of mentions in a given paragraph, one of the points will cover the other, obscuring the data. The jitter eliminates this problem.


# Question 10
Do Harry and Hermione tend to co-occur?  Fit a linear regression of Harry vs. Hermione references per paragraph. Use the `lm()` function and print out the `summary` of the model.


```{r}
lm(Hermione ~ Harry, paragraph_df) %>%
  summary()
```


Now use `geom_smooth` to plot the linear regression line on top of the jitter plot.

```{r}
ggplot(paragraph_df, aes(x = Harry, y = Hermione)) + 
  geom_jitter(width = .3, height = .3) +
  geom_smooth(method = 'lm')
```


# Question 11
Is there are relationship between the length of the paragraph a the number of times Harry is mentioned? Add a column called `num_words` to `paragraph_df` counting the number of words in each paragraph. Then use a linear regression to answer for the question. Provide both a statistical summary and a visualization.


```{r}
paragraph_df <- paragraph_df %>%
  mutate(num_words = str_count(text, "\\w+"))
lm(Harry ~ num_words, paragraph_df) %>%
  summary()
ggplot(paragraph_df, aes(x = num_words, y = Harry)) + 
  geom_point() +
  geom_smooth()

```

# Question 12
Create an indicator variable `harry_mentioned` that indicates whether or not Harry is mentioned in each paragraph. This indicator variable should be a factor (e.g. use `as.factor`).

```{r}
paragraph_df <- paragraph_df %>%
  mutate(harry_mentioned = as.factor(Harry > 0))
paragraph_df
person_refs
```

Now repeat the previous linear regression with `harry_mentioned` as the x variable instead of the number of times he is mentioned

```{r}
lm(num_words ~ harry_mentioned, paragraph_df) %>%
  summary()
ggplot(paragraph_df, aes(x = harry_mentioned, y = num_words)) + 
  geom_jitter(width = .2) +
  geom_smooth()
```


# Free response
Ask and answer a question with this data set. You should make at least 2 figures (e.g. plot, printout of a regression, etc). Provide a written explanation of the question and the evidence for your answer.


I'd like to see if the presence of Ron along with Harry in a paragraph is a predictor of the presense of Hermione

``` {r}
paragraph_df <- paragraph_df %>%
  mutate(harry_ron_mentioned = as.factor(Harry > 0 & Ron > 0), 
         hermione_mentioned = as.factor(Hermione > 0)) %>%
  filter()

ggplot(paragraph_df, aes(x = harry_ron_mentioned, y = Hermione)) + 
  geom_jitter(width = .2, height = .3) +
  geom_smooth()

lm(Hermione ~ Harry + Ron, filter(paragraph_df, Harry > 0 & Ron > 0)) %>%
  summary()
```





